Hippocampal segmentation with MAGeT

### 2013-01-24 paper.Rnw
Full incorporated comments from Mallar into the introduction

In this paper we propose an extension to the basic multi-atlas-based
segmentation scheme to address this issue.  Principally, we explore the
possibility of using a small atlas library to bootstrap a much larger
\emph{template library} composed of images taken from the target population.
The template library is then used to segment the subjects in a similar fashion
to basic multi-atlas segmentation: by label propagation and label fusion.  This
intuition driving this approach is to leverage the unique neuroanatomy of
subject population on hand to initialize the segmentation process and improve
accuracy over direct propagation from the atlas library to unlabelled subjects.

The essential insight of generating a template library is not new.  Heckemann
\citep{Heckemann2006} briefly explored "indirect" segmentation by taking a
single atlas and propagating those labels to intermediate subjects before
fusing them in a target image space.  Heckemann compared indirect single-atlas
segmentation to direct multi-atlas segmentation found the indirect approach
performed worse.  In this paper we continue the same line of investigation but
explore the performance when using multiple atlases, as well as several
different registration and fusion methods.

@@

In previous work \citep{Chakravarty2012}, we applied MAGeT brain to
segmentation of the human striatum, globus pallidus, and thalamus using a
single histologically-derived atlas. The main contribution of this paper is to
extend our approach to the human hippocampus - a much more difficult structure
to segment -- and perform a thorough validation over a range of atlas and
template library sizes, registration methods, and fusion methods.  Our aim in
this work is not necessarily to improve on segmentation accuracy beyond
existing methods, but instead to provide a method that trades off manual
segmentation expertise and time for computational processing time while still
providing sufficient accuracy for clinical and research applications.

>>>>

In this paper we propose an extension to the basic multi-atlas-based
segmentation scheme to address this issue.  Principally, we explore the
possibility of using a small atlas library to bootstrap a much larger
\emph{template library} composed of images taken from the target population.
The template library is then used to segment the subjects in a similar fashion
to basic multi-atlas segmentation: by label propagation and label fusion.  We
call our approach MAGeT brain: {\em m}ultiple {\em a}utomatically {\em
ge}nerated {\em t}emplate {\em brain} segmentation. The intuition driving this
approach is that by generating a template library we leverage the unique
neuroanatomy of subject population on hand to initialize the segmentation
process and improve accuracy over direct propagation from the atlas library to
unlabelled subjects while also using fewer manually segmented atlases.

The insight of generating a template library is not new.  Heckemann
\citep{Heckemann2006} briefly explored "indirect" segmentation by taking a
single atlas and propagating those labels to intermediate subjects before fusing
them in a target image space.  Heckemann compared indirect single-atlas
segmentation to direct multi-atlas segmentation found the indirect approach
performed worse.  In this paper we continue the same line of investigation but
explore the performance when using multiple atlases, as well as several
different registration and fusion methods.

@@

In previous work \citep{Chakravarty2012}, we applied MAGeT brain to segmentation
of the human striatum, globus pallidus, and thalamus using a single
histologically-derived atlas. The contribution of the present work is to extend
our approach to the human hippocampus and perform a rigorous validation. We
first conduct an intensive cross-validation of MAGeT on a pool of 69 MRIs from
the ADNI database, comparing the performance of MAGeT against basic multi-atlas
segmentation using a range of atlas and template library sizes, registration and
fusion methods.  We then compare MAGeT segmentations of the entire ADNI
screening dataset with those of established automated and manual segmentation
techniques.  Finally, we compare MAGeT segmentations to those of manual
segmentations from a dataset of first episode schizophrenic patient MRIs.
